---
title: "Barrel percent by launch angle"
author: "Nick Hatley"
date: "2019-09-04"
output:
  blogdown::html_page
categories: ["R"]
slug: barrels

tags: ["Statcast", "Exit Velocity", "Barrels"]
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning =  FALSE)
knitr::opts_chunk$set(message = FALSE)
options(knitr.table.format = "html")
```

```{r get librarys, function and data}
library(tidyverse)
mlb_db_dir = "~/Dropbox/mlb_db/"

source(paste0(mlb_db_dir, "code/functions/plot_theme.R"))

barrels_15_19 <- map(2015:2019,
                     ~baseballr::scrape_savant_leaderboards(
                       leaderboard = "exit_velocity_barrels", 
                       year = .x)) %>% 
  bind_rows

```

Data from this post retreived via the fabulous [`baseballr`](http://billpetti.github.io/baseballr/) package from Bill Petti.   

It has a number of data acquisition and other functions. Here, I used the `baseballr::scrape_savant_leaderboards()` function to fetch data from the MLB's [`baseball savant`](https://baseballsavant.mlb.com) website for exit velocity as measured by [`Statcast`](http://m.mlb.com/statcast/leaderboard#exit-velo,r,2019).  

**********  


The resulting dataset looks like:  
```{r, echo = TRUE, collapse = TRUE, out.width='300%'}
barrels_15_19 %>% 
  glimpse()
```

  
**********     
     
Code is at the end of the post.   

**********  

```{r, fig.height=8.5, fig.width=10.5, fig.align='center'}
p = barrels_15_19 %>% 
  group_by(last_name, first_name, player_id) %>% 
  summarise(attempts = sum(attempts),
            barrels = sum(barrels),
            seasons = n(),
            avg_hit_speed = sum(avg_hit_speed) / seasons,
            avg_distance = sum(avg_distance) / seasons,  
            avg_hr_distance = sum(avg_hr_distance) / seasons,
            brl_percent = sum(brl_percent) / seasons,
            avg_hit_angle = sum(avg_hit_angle) / seasons,
            anglesweetspotpercent = sum(anglesweetspotpercent) / seasons
            ) %>% 
  ungroup %>% 
  mutate(player = paste(first_name, last_name, sep = " ")) %>% 
  filter(avg_hit_angle > 0) %>% 
  ggplot(
    aes(
      x = brl_percent, 
      y = avg_hit_angle, 
      group = player
    )
    ) + 
  geom_jitter(
    shape = 21, stroke = rel(1.25),
    colour = "#13274F", fill = "white"
  ) + 
  ggrepel::geom_text_repel(data = . %>% 
                             filter((avg_hit_angle < 20 & brl_percent > 16) |
                                      (avg_hit_angle > 20 & brl_percent > 10)),
                          aes(
                            label = player
                          )
                           ) + 
  plot_theme() + 
  labs(x = "Barrel Percent",
       y = "Average Launch Angle for Hits"
       ) 

plot(p)
```

************
  
Interactive version:    
  
```{r, fig.height=8.5, fig.width=10.5}
p_ly = plotly::ggplotly(p)

p_ly
```


*********

Code for post:
```{r, eval=FALSE, echo = TRUE}
##get data via 
library(tidyverse)

barrels_15_19 <- map(2015:2019,
                     ~baseballr::scrape_savant_leaderboards(
                       leaderboard = "exit_velocity_barrels", 
                       year = .x)) %>% 
  bind_rows

##ggplot plot
p = barrels_15_19 %>% 
  group_by(last_name, first_name, player_id) %>% 
  summarise(attempts = sum(attempts),
            barrels = sum(barrels),
            seasons = n(),
            avg_hit_speed = sum(avg_hit_speed) / seasons,
            avg_distance = sum(avg_distance) / seasons,  
            avg_hr_distance = sum(avg_hr_distance) / seasons,
            brl_percent = sum(brl_percent) / seasons,
            avg_hit_angle = sum(avg_hit_angle) / seasons,
            anglesweetspotpercent = sum(anglesweetspotpercent) / seasons
            ) %>% 
  ungroup %>% 
  mutate(player = paste(first_name, last_name, sep = " ")) %>% 
  filter(avg_hit_angle > 0) %>% 
  ggplot(
    aes(
      x = brl_percent, 
      y = avg_hit_angle, 
      group = player
    )
    ) + 
  geom_jitter(
    shape = 21, stroke = rel(1.25),
    colour = "#13274F", fill = "white"
  ) + 
  ggrepel::geom_text_repel(data = . %>% 
                             filter((avg_hit_angle < 20 & brl_percent > 16) | (avg_hit_angle > 20 & brl_percent > 10)),
                          aes(
                            label = player
                          )
                           ) + 
  plot_theme() + ##this is just farily simple plot theme I made
  labs(x = "Barrel Percent",
       y = "Average Launch Angle for Hits"
       ) 

##plotly interactive plot
p_ly = plotly::ggplotly(p)

```

   
   
   