---
title: "Crpytocurrency Charts"
author: "Nick Hatley"
date: "`r format(Sys.Date())`"
output:
  blogdown::html_page
categories: ["R"]
slug: cryptocurrency-charts

tags: ["R Markdown", "blogdown"]
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning =  FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.align = 'center')
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(fig.height = 7.5)
options(knitr.table.format = "html")
```


```{r get librarys, function and data}
library(tidyverse)
library(plotly)
library(knitr)
library(htmltools)
library(kableExtra)

bitstamp_btc_usd = "BITSTAMP_SPOT_BTC_USD"
bitstamp_eth_usd = "BITSTAMP_SPOT_ETH_USD"
bitstamp_ltc_usd = "BITSTAMP_SPOT_LTC_USD"
bitstamp_xrp_usd = "BITSTAMP_SPOT_XRP_USD"
bitstamp_bch_usd = "BITSTAMP_SPOT_BCH_USD"

coins = list(bitstamp_btc_usd, bitstamp_eth_usd, bitstamp_ltc_usd,
             bitstamp_xrp_usd, bitstamp_bch_usd)

all_data = read_rds(here::here("data", "all_data.rds"))

make_ohlc_candlestick_md <- function(symbol_in, time_frame) {
    df2 = all_data %>%
    filter(duration==time_frame) %>%
    filter(symbol==symbol_in)

    p = df2 %>%
    plotly::plot_ly(x = ~time_open, y = ~price_close, type="candlestick",
            open = ~price_open, close = ~price_close,
            high = ~price_high, low = ~price_low) %>%
    plotly::add_lines(x = ~time_open, y = ~price_close,
              line = list(color = 'black', width = 0.75), inherit = F) %>%
    plotly::layout(showlegend = FALSE) %>%
    plotly::layout(title = paste0(symbol_in))

}
```
If you aren't keen on finance or if you are new to these kinds of charts, as I was, I'll explain them briefly:

>Each individual bars represent a day. The low point of the bars represent the low-point for the coin that day and so forth for the high/high point. The coloring represents where the coin closed that day in respect to the prior day. Red = below, Green = above.

The charts are interactive and they include a 'rangeslider' on the bottom that allows you to zoom into specific time-spans.  

Data from this post was retrieved on:
```{r}
all_data$time_period_end[[1]]
```



### Six months {.tabset}
```{r, six_months}
six_months = map(coins, function(x)
  make_ohlc_candlestick_md(symbol = x, time_frame = "six_months"))
htmltools::tagList(six_months)
```

***********

### One year {.tabset}
```{r, one_year}
one_year = map(coins, function(x)
  make_ohlc_candlestick_md(symbol = x, time_frame = "one_year"))
htmltools::tagList(one_year)
```

***********

### Two years {.tabset}
```{r, two_years}
two_years = map(coins, function(x)
  make_ohlc_candlestick_md(symbol = x, time_frame = "two_years"))
htmltools::tagList(two_years)
```
