---
title: "Nick Markakis' Defense is Really Bad"
author: "Nick Hatley"
date: "2019-11-25"
output:
  blogdown::html_page
categories: ["R"]
slug: markakis_defense

tags: ["mlb", "defense"]
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning =  FALSE)
knitr::opts_chunk$set(message = FALSE)
options(knitr.table.format = "html")
```


```{r get librarys, function and data}
library(tidyverse)
mlb_db_dir = "~/Dropbox/mlb/"
outfielders <- read_rds(paste0(mlb_db_dir, "output/outfielders_value.rds"))
source(paste0(mlb_db_dir, "code/functions/plot_theme.R"))

```

Nick Markakis' defense is **really, really** bad.   
  
That's it, that's the post.

  
*************  
```{r, fig.height=14, fig.width=12.5, fig.align='center'}

outfielders %>% 
  mutate(markakis = name == "Nick Markakis") %>% 
  mutate(first_initial = str_extract(first_name, "^..")) %>% 
  mutate(name_label = paste0(first_initial, ". ", last_name)) %>%  
  ggplot(
    aes(
      x = reorder(name, def_per_pa),
      y = def_per_pa
    )
  ) +
  geom_col(
    aes(
      x = reorder(name, def_per_pa),
      y = def_per_pa,
      fill = markakis
      )
  ) + 
  geom_segment(
      data = . %>% filter(name == "Nick Markakis"),
        aes(
          x = 30,
          xend = name,
          y = .01,
          yend = .0001
        ),
      arrow = arrow(length = unit(0.03, "npc"))
  ) + 
  annotate("text",
           x = 33, 
           y = .01, 
           label =
             "Markakis had the 8th worst\noutfield defensive value last year") +
  coord_flip() + 
  plot_theme() + 
  labs(y = "Defensive Value per Plate Attempt") + 
  ggtitle("Nick Markakis' defense was really really bad in 2019", 
          subtitle = "Among OF with >200 plate attempts") + 
  theme(legend.position = 'none',
        axis.title.y = element_blank(), 
        axis.text.y = element_text(size = rel(.85)),
        plot.subtitle = element_text(face = "italic"),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank()
  )   
  
  
```

***********
  
  
A couple things stand out to me. Take the Astros and Yankees. The Astros have a hump at around 80 or so wrc+, but over half their team is >100 and is pretty evenly distributed up to the 150 or so range. The number of >150 wrc+ hitters in their lineup is unheard of across the league and is a big reason why they lead the league with the team average of 122. The Yankees on, the other hand, have almost no 150 or higher wrc+ hitters, but almost all of their hitters are above league average.   
  
  
***********
```{r, out.width = "60%", out.height= "40%", fig.align='center'}
p = outfielders %>% 
  mutate(markakis = name == "Nick Markakis") %>% 
  
  ggplot(
    aes(
      x = off,
      y = def
    )
  ) +
  geom_point(
    aes(
    text = name,
    color = markakis
    )
  ) +
  geom_vline(xintercept = 0) + 
  geom_hline(yintercept = 0) + 
  theme_bw() + 
  theme(legend.position = 'none') + 
  labs(x = "Offensive Value",
       y = "Defensive Value"
       )


p_ly = plotly::ggplotly(p, tooltip = c("text", "x", "y"), width = 1300, height = 550)

p_ly
```

***********  

